---
title: "Replication"
author: "Econ 771: Health Economics II"
date: "Due: Friday, December 11, 2020"
output: 
  bookdown::html_document2:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RefManageR)
source('paths.R')
bib <- ReadBib(paste0(bib.path,'/BibTeX_Library.bib'))
myopts <- BibOptions(bib.style = "authoryear", style="citation", first.inits=FALSE, max.names = 20, check.entries=FALSE)

```



# Overview
This describes the logistics and expectations for our main class project, which is a replication of an empirical paper in health economics. Note that by "replication", I don't necessarily mean a 1-to-1 match in the data and empirical anlalysis. That's possible in some settings where the authors use entirely publicly available and easily accessible data, but more often than not, authors use a combination of datasets. Some of those datasets are restricted, and some are technically publicly available but require a ton of cleanup to use. I want us to get *some* experience working with real-life data, but I don't want it to take months of your time. So...the expectation is that you make an attempt to replicate the authors' identification strategy and econometric analysis for at least a subset of the questions/analyses addressed in the paper. Essentially, you're going to work with some real data and use the paper as a guide for how to go about your analysis.


# Potential papers
Here is a list of paper's that I think are sufficiently replicable using publicly available data. 

```{r, echo = FALSE, results="asis"} 
bib[c("duggan2000","horwitz2009","dafny2005","dranove2019","dranove2017","geruso2020")]
``` 


No paper on this list is entirely replicable because they all use a combination of different datasets. There's a more detailed rubric in the next section where I describe the expectations of your replication exercise. Here, I just note how you can go about the replication for any given paper:

- Duggan (2000): This is replicable with historic IPPS data that allow you to proxy for a hospital's share of Medicaid or uninsured patients.
- Horwitz and Nichols (2009): This is largely reproducible with data from the American Hospital Association (AHA). We have access to these data through the Wharton Research Data Center, but there are some restrictions. So we'll just need to chat first before you proceed with this paper as your replication choice.
- Dafny (2005): You can use the case mix index from the IPPS files to proxy for DRG coding. It's not perfect, but again, the point is to work with read data using a published paper as your guide.
- Dranove and Ody (2019): This is also doable if you use a different measure of physician practice ownership from the AHA surveys.
- Dranove and Ody (2017): Reproducible with the Hospital Cost Report data.
- Geruso and Layton (2020): Section 5 of this paper uses publicly available Medicare Advantage data, so that's fully reproducible. In other analyses, the authors use Massachusetts All Payer Claims Data, which is also available but may be difficult to work with in a semester. Nonetheless, if you think those data may be useful for other purposes, then it could be worth it to start that investment in this class.



# Advice
