---
title: "Replication"
author: "Econ 771: Health Economics II"
date: "Due: Friday, December 11, 2020"
output: 
  bookdown::html_document2:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    theme: darkly
    number_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RefManageR)
source('paths.R')
bib <- ReadBib(paste0(bib.path,'/BibTeX_Library.bib'))
myopts <- BibOptions(bib.style = "authoryear", style="html", first.inits=FALSE, max.names = 20, check.entries=FALSE)

```



# Overview
This describes the logistics and expectations for our main class project, which is a replication of an empirical paper in health economics. Note that by "replication", I don't necessarily mean a 1-to-1 match in the data and empirical anlalysis. That's possible in some settings where the authors use entirely publicly available and easily accessible data, but more often than not, authors use a combination of datasets. Some of those datasets are restricted, and some are technically publicly available but require a ton of cleanup to use. I want us to get *some* experience working with real-life data, but I don't want it to take months of your time. So...the expectation is that you make an attempt to replicate the authors' identification strategy and econometric analysis for at least a subset of the questions/analyses addressed in the paper. Essentially, you're going to work with some real data and use the paper as a guide for how to go about your analysis.

To begin, please "accept" the assignment on our GitHub classroom page. This will initialize a GitHub repository where you can edit and ultimately craft your replication in your data management/econometrics software of choice.  

You will "turn in" your repository to me with your replication and all supporting documents. Please create a final document with your main results in a PDF. Please also be sure to include in your repository all of your supporting code files. Practice writing good code and showing me only what I would need to recreate your results.

You have until the final exam date to submit your replication (**December 11, 2020**).


# Potential papers
Here is a list of papers that I think are sufficiently replicable using publicly available data. No paper on this list is entirely replicable because they all use a combination of different datasets. There's a more detailed rubric in the next section where I describe the expectations of your replication exercise. Here, I just note how you can go about the replication for any given paper. These papers are listed in order of appearance in the class schedule. 

You are also welcome to select your own paper, but you need to be sure that it is replicable. Sometimes papers seems easy to replicate but end up being impossible due to some specific proprietary data they have.


---

```{r, echo = FALSE, results="asis"} 
bib["duggan2000"]
``` 

This is replicable with historic IPPS data that allow you to proxy for a hospital's share of Medicaid or uninsured patients; however, data on hospital disproportionate share isn't readily available before 1994. That's the first year the variable is populated in the historic IPPS data.

---

```{r, echo = FALSE, results="asis"} 
bib["horwitz2009"]
``` 

This is largely reproducible with data from the American Hospital Association (AHA). We have access to these data through the Wharton Research Data Center, but there are some restrictions. So we'll just need to chat first before you proceed with this paper as your replication choice.


---

```{r, echo = FALSE, results="asis"} 
bib["dranove2019"]
``` 

This is also doable if you use a different measure of physician practice ownership from the AHA surveys. For access to the AHA data, just get in touch with me and we can try to get you set up with an Emory account.


---

```{r, echo = FALSE, results="asis"} 
bib["dranove2017"]
``` 

Reproducible with the Hospital Cost Report data, [HCRIS](https://github.com/imccart/HCRIS).


---

```{r, echo = FALSE, results="asis"} 
bib["geruso2020"]
``` 

Section 5 of this paper uses publicly available Medicare Advantage data, so that's fully reproducible. In other analyses, the authors use Massachusetts All Payer Claims Data, which is also available but may be difficult to work with in a semester. Nonetheless, if you think those data may be useful for other purposes, then it could be worth it to start that investment in this class.


---

```{r, echo = FALSE, results="asis"} 
bib["molitor2018"]
``` 

Parts of this paper are approximately replicable using the count of claims based on HCPCS codes and NPIs. You can also identify physician movers using the physician compare data. The outcomes will differ slightly but this approach would be analogous to the paper.


---

```{r, echo = FALSE, results="asis"} 
bib["gruber1996"]
``` 

Most of this analysis should be replicable using the [National Hospital Discharge Survey](https://www.cdc.gov/nchs/nhds/index.htm).


---

```{r, echo = FALSE, results="asis"} 
bib["clemens2017"]
``` 

Analysis of physician practices (Section IV.E of the paper) is replicable using the [Community Tracking Study](https://www.icpsr.umich.edu/web/ICPSR/series/161). You can also reconstruct physician payments using the Physician Fee Schedule data.



---

```{r, echo = FALSE, results="asis"} 
bib["baicker2013"]
``` 

You can approximately replicate this paper using a combination of data from the [Area Health Resources File](https://data.hrsa.gov/topics/health-workforce/ahrf) and MA penetration and benchmark data. The MA data are available on my GitHub repo, [Medicare Advantage Repository](https://github.com/imccart/Medicare-Advantage). You could also mimic many of the cost outcomes using the [HCRIS](https://github.com/imccart/HCRIS) data, which are also part of a GitHub repo.


---

```{r, echo = FALSE, results="asis"} 
bib["ericson2014"]
``` 

You can replicate this paper using data from Medicare Part D. These raw data are available from CMS and the clean data are available from the AEJ: Policy website, [Ericson (2014)](https://www.aeaweb.org/articles?id=10.1257/pol.6.1.38). If you choose this paper as your replication, you must download the data from its raw source. [CMS Part D Enrollment](https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/MCRAdvPartDEnrolData) is a good starting point, but you'll probably have to navigate the website to find all of the other variables.


---

```{r, echo = FALSE, results="asis"} 
bib["dafny2008"]
``` 

You can "replicate" this paper using more recent data on Medicare Advantage and the star rating program. This won't be the same analysis but it will provide good practice with the MA data, which is accessible from the [Medicare Advantage Repository](https://github.com/imccart/Medicare-Advantage).


---

```{r, echo = FALSE, results="asis"} 
bib["dranove2016"]
``` 

You can approximately replicate this paper using data from the Hospital Cost Reports, [HCRIS](https://github.com/imccart/HCRIS), and [Provider of Services](https://github.com/asacarny/provider-of-services) files. For information on how to define hospital markets, be sure to take a look at this [healthcare markets GitHub repo](https://github.com/graveja0/health-care-markets).


---

```{r, echo = FALSE, results="asis"} 
bib["cooper2017"]
```

You can't replicate this analysis directly because they use commercial insurance data; however, you can recreate most of their analysis using an estimate of price from [HCRIS](https://github.com/imccart/HCRIS).


---

```{r, echo = FALSE, results="asis"} 
bib["dafny2009"]
```

You can replicate this analysis using [HCRIS](https://github.com/imccart/HCRIS) and the [Provider of Services](https://github.com/asacarny/provider-of-services) files. You will also need to track down data on hospital mergers and acquisitions. As an Emory student, you all have access to the AHA annual survey data which includes information that you can use to identify mergers. A better data source is from the [Healthcare Pricing Project](https://healthcarepricingproject.org/), which provides merger data from 2001 through 2014 (based on AHA surveys).


---

```{r, echo = FALSE, results="asis"} 
bib["dafny2019"]
```

Most of this analysis is replicable using [HCRIS](https://github.com/imccart/HCRIS) and the [Provider of Services](https://github.com/asacarny/provider-of-services) files, along with the merger data from the [Healthcare Pricing Project](https://healthcarepricingproject.org/).



---

```{r, echo = FALSE, results="asis"} 
bib["schmitt2018"]
```

Most of this analysis is replicable using [HCRIS](https://github.com/imccart/HCRIS) and the [Provider of Services](https://github.com/asacarny/provider-of-services) files, along with the merger data from the [Healthcare Pricing Project](https://healthcarepricingproject.org/). But a word of warning...measuring multi-market contact can be computationally cumbersome.


# Expectations
Replicating empirical work is really hard and always underestimated in terms of the time and effort necessary to do it. I **do not** expect that you can match someone's results exactly (or even approximately). The point is to try and see where your results differ. 

To that end, you should state clearly the elements of the paper and research quetions(s) you're trying to replicate. Describe your data. Present your estimation strategy and results. And ultimately discuss your data and results relative to the paper you're replicating.

# Grading
Your grade will be based on 5 items, each worth 20 points toward the final assignment grade. For each component, you'll receive a 0 (not present), 10 (the element is there but poorly executed), or 20 (element is there and sufficiently executed). The 5 elements are listed below:

1. What parts of the paper are you attempting to replicate?
2. What data are you using?
3. What econometric tools are you using (identification strategy, estimator, and estimating equation)?
4. What are your results?
5. How do your data and results differ from the paper you're replicating? Why? You may not really know why your results differ, but you should at least be able to describe some potential reasons.