\documentclass[11pt,]{article}
\usepackage[margin=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}
\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center
\newcommand{\blankline}{\quad\pagebreak[2]}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}} 
\usepackage{longtable,booktabs}

\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}

\titleformat*{\subsubsection}{\normalsize\itshape}

\usepackage{titling}
\setlength{\droptitle}{-.25cm}

%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines 

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{lastpage}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0.0pt} 
\lhead{}
\chead{}
\rhead{\footnotesize Econ 771: Health Economics II -- 8/19/2020 -
12/11/2020}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\fancypagestyle{firststyle}
{
\renewcommand{\headrulewidth}{0pt}%
   \fancyhf{}
   \fancyfoot[C]{\small \thepage/\pageref*{LastPage}}
}

%\def\labelitemi{--}
%\usepackage{enumitem}
%\setitemize[0]{leftmargin=25pt}
%\setenumerate[0]{leftmargin=25pt}




\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
}
\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={Econ 771: Health Economics II},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls


\setcounter{secnumdepth}{0}





\usepackage{setspace}

\title{Econ 771: Health Economics II}
\author{Ian McCarthy}
\date{8/19/2020 - 12/11/2020}


\begin{document}  

		\maketitle
		
	
		\thispagestyle{firststyle}

%	\thispagestyle{empty}


	\noindent \begin{tabular*}{\textwidth}{ @{\extracolsep{\fill}} lr @{\extracolsep{\fill}}}


E-mail: \texttt{\href{mailto:ian.mccarthy@emory.edu}{\nolinkurl{ian.mccarthy@emory.edu}}} & Web: \href{http://ianmccathyecon.com}{\tt ianmccathyecon.com}\\
Office Hours: W 09:30-11:30 a.m.  &  Class Hours: TTh 9:40-11:30 a.m.\\
Office: 319 Rich Memorial Building  & Class Room: \emph{online}\\
	&  \\
	\hline
	\end{tabular*}
	
\vspace{2mm}
	


\hypertarget{course-description}{%
\section{Course Description}\label{course-description}}

This course explores the industrial organization of healthcare markets
in the U.S. We will focus on the following areas: hospital ownership and
production, physician agency, quality disclosure, and hospital
competition. If we have time, we will briefly cover some issues on
health insurance markets. The class is effectively designed as a
combination of empirical IO and causal inference, with applications to
healthcare. As such, we will also examine several econometric tools and
causal inference identification strategies. These methods will be
introduced as needed throughout the course.

Supply-side health economics is a rapidly growing field with many new
developments, particularly in the areas of bargaining in two-sided
markets and physician learning. Some of these recent developments use
tools from network analysis and machine learning, which we unfortunately
do not have time to cover in this course. I've also chosen specific
topics that overlap most with my own research --- the upside here is
that I can speak somewhat confidently about the literature and empirical
studies in this area, but the downside is that some very interesting
areas of health economics are not studied. For example, we will largely
ignore issues of the prescription drug market, medical devices, and
physician labor supply. My hope is that the content that we do cover
will provide a springboard for those interested in these other important
areas.

\hypertarget{learning-outcomes}{%
\section{Learning Outcomes}\label{learning-outcomes}}

I have four central goals for this course:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Synthesize the current literature in each of the main areas of health
  economics covered in this class
\item
  Apply standard causal inference techniques in the area of healthcare
\item
  Provide constructive criticism of academic work in this area
\item
  Develop your own preliminary research in some area of health economics
\end{enumerate}

There are more specific learning outcomes for each module described on
the relevant module's page of our class website, accessible
\href{https://imccart.github.io/Econ-771/}{here}.

\hypertarget{text-software-and-class-materials}{%
\section{Text, Software, and Class
Materials}\label{text-software-and-class-materials}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Readings:} As an elective PhD course, we will rely on academic
  papers from the reading list below. I expect everyone to read the
  papers in advance and come to class with questions on the study's
  contribution, empirical techniques, identification strategies, and
  datasets used. Where possible, we will work to replicate results or
  apply the paper's central methodology to some simulated data. My goal
  with each paper is to discuss the analysis in detail. As such, the
  primary reading list is perhaps shorter than a standard PhD course.
  I've provided supplemental reading in each section for those
  interested in additional readings in a specific area.
\item
  \textbf{Statistics Software:} For any data analysis or econometrics,
  I'll use \texttt{R} for my work, but you are free to use whatever
  software you're most comfortable with. I encourage you to use
  \texttt{R}, \texttt{Stata}, or \texttt{Python} simply because these
  are the most common programs used in practice right now.
\item
  \textbf{Accessing Data:} For any in-class activities, I will house all
  of the code and links to the data on our class website. For any
  real-life datasets, I will also point you to other GitHub repositories
  where available. So you'll need to be sure to have a simple working
  knowledge of Git and GitHub. If you're new to Git or GitHub, take a
  look at \href{https://grantmcdermott.com/}{Grant McDermott's} notes on
  \href{https://github.com/uo-ec607/lectures}{Data Science for
  Economists} as well as \href{https://jennybryan.org/}{Jenny Bryan's}
  online reference book, \href{https://happygitwithr.com/}{Happy Git and
  GitHub with R}.
\item
  \textbf{Cloud Storage and Analysis:} You have access Amazon Web
  Services (AWS) Cloud Computing services for all of the empirical work
  in the class through the AWS Classroom. This \emph{should} act as a
  virtual console that each of you can access independently but where
  you are each working on an identical system. This avoids issues of
  different versions of stats programs or, in the case of \texttt{R} and
  \texttt{Python}, different packages or package versions. Using AWS
  isn't required for the class, but it is available to you for those
  interested.
\item
  \textbf{Logistics:} For day-to-day communication, grades, and other
  private information (such as Zoom meeting links), I'll use
  \emph{Canvas}. I'll post all other materials to our class website.
\item
  \textbf{Lectures and Notes:} Any presentations will be made available
  on our class website prior to any given class. I may also annotate
  those slides during each class, for which I'll use \emph{OneNote}.
  Everyone has access to this as an Emory student, so I will set up a
  classroom notebook which will serve as a single resource for all class
  notes. You can also take your own notes as part of the app, which
  should help with keeping things organized.
\item
  \textbf{Online Discussions:} We'll use
  \href{https://grouptweet.com/}{GroupTweet} to facilitate online
  discussions this semester. This will work with any smartphone. It's
  basically our own private Twittersphere for our class. You will
  receive an invitation to join the group, at which point we can post
  and respond throughout the semester. Participation in these online
  discussions is a small part of your grade, as discussed below.
\item
  \textbf{Turning things in:} We'll use GitHub Classroom to facilitate
  all of the assignments. You'll soon receive an invite which will get
  you access to the classroom. Within the classroom, you will have to
  accept each assignment, which will then create a separate GitHub
  repository for you for each assignment. You can then do all of your
  work in whatever software program you'd like, and then commit and push
  your changes to your repository. I will see this repository and can
  comment on it, which is how I will provide feedback on each
  assignment.
\end{enumerate}

\hypertarget{course-policies}{%
\section{Course Policies}\label{course-policies}}

Various policies for this course are described below. Basically, let's
all work to be good citizens and take seriously our various roles as a
student, teacher, friend, colleague, human, etc.

\hypertarget{academic-integrity}{%
\subsection{Academic integrity}\label{academic-integrity}}

The Emory University Honor Code is taken seriously and governs all work
in this course. Details about the Honor Code are available in the Laney
Graduate School Handbook and available online
\href{https://www.gs.emory.edu/academics/policies-progress/conduct.html}{here}.
By taking this course, you affirm that it is a violation of the code to
plagiarize, to deviate from the instructions about collaboration on work
that is submitted for grades, to give false information to a faculty
member, and to undertake any other form of academic misconduct. You also
affirm that if you witness others violating the code you have a duty to
report them to the honor council.

\hypertarget{accessibility-services}{%
\subsection{Accessibility services}\label{accessibility-services}}

If you anticipate issues related to the format or requirements of this
course, please meet with me. I would like us to discuss ways to ensure
your full participation in the course. If you determine that
accommodations are necessary, you may register with Accessibility
Services at (404)727-9877 or via e-mail at
\href{mailto:accessibility@emory.edu}{\nolinkurl{accessibility@emory.edu}}.
To register with OAS, students must self identify and initiate contact
with the OAS office.

\hypertarget{communication}{%
\subsection{Communication}\label{communication}}

I will post regular announcements to the class on \emph{Canvas}, so
please set up your notifications on \emph{Canvas} accordingly. I will
also use \emph{Canvas} to post all grades and any other information that
needs to stay in the class (like our \emph{Zoom} meeting link for
virtual meetings). All other course materials will be available on our
class website,
\href{https://imccart.github.io/Econ-771/}{imccart.github.io/Econ-771/}.

Please feel free to reach out to me for any reason. I generally respond
to all e-mails within 24 hours.

\hypertarget{virtual-meetings}{%
\subsection{Virtual meetings}\label{virtual-meetings}}

All meetings and office hours will take place virtually this semester on
\emph{Zoom}. I'll use the same \emph{Zoom} meeting room for all
meetings, the details of which are available on our \emph{Canvas} site.
If you can't make it to my usual office hours, please feel free to
schedule another time to meet (virtually). Just follow the link to
select a time that works for you,
\url{https://ian-meetings.youcanbook.me}.

\hypertarget{attendance}{%
\subsection{Attendance}\label{attendance}}

While there is no official ``attendance'' credit, everyone is expect to
attend all class sessions. Given our small class, it is very important
that we are all present and engaged.

\hypertarget{assignments-and-grading}{%
\section{Assignments and Grading}\label{assignments-and-grading}}

There are four main assignments throughout the semester, along with a
participation component. I describe each assignment below, with much
more detail provided on our
\href{https://imccart.github.io/Econ-771/}{class website}.

\hypertarget{participation}{%
\subsection{Participation}\label{participation}}

We will have semi-weekly online discussion forums in GroupTweet where we
discuss a variety of issues, ranging from specific areas of confusion on
some papers to best coding practices for empirical work. I anticipate
close to one discussion forum per week. Your final participation grade
will reflect your percentage (out of a maximum of 10) of discussion
forums in which you participate.

\hypertarget{empirical-exercises}{%
\subsection{Empirical exercises}\label{empirical-exercises}}

Each course module will have an applied component where we spend some
time replicating analyses from selected papers based on real-world data.
These will require some of your time outside of class to get the data in
working order and implement the relevant identification strategy and
econometric estimator. This empirical work should heavily complement
your replication project as well as your draft paper. The due dates for
all empirical exercises are indicated in the schedule below. In general,
we have one empirical assignment due at the end of each module.

\hypertarget{presentations}{%
\subsection{Presentations}\label{presentations}}

You will present two papers throughout the course of the semester. A
list of potential papers will be provided on our class website within
the first week of class. Please inform me of your selected papers no
later than \textbf{August 21, 2020}.

Each presentation should be no more than 30 minutes, including questions
and class discussion. The presentation should follow a standard
conference setup, with a brief introduction/motivation, a very brief
discussion of the literature and some context of the paper, discussion
of the data, empirical analysis, and results.

Note that a presentation is not just a re-hashing of the paper in slide
form. A good academic presentation should have as little information as
possible on each slide, and the content on the slides doesn't
necessarily need to follow that of the paper. For example, in a
real-time environment, it is much easier to move between different
aspects of the empirical analysis and data.

In addition to the presentation, please send me your slides in advance.
For people in the economics department, I expect the slides to be
completed in Beamer (LaTeX), \texttt{R\ Markdown}, or Python. Others can
use PowerPoint or some other presentation tool as they see fit. My only
recommendation with those other programs is that, for some of them like
Prezi or Powtoon, it is easy for the presentation to become distracting.
Slides should complement your presentation and not replace your
presence!

Grades for each presentation will be based 50\% on the slides and 50\%
on the actual presentation (delivery, clarity/organization, and
content). Additional details and a grading rubric are available on the
assignments page in our class website.

\hypertarget{literature-review-2nd-year-students}{%
\subsection{Literature review (2nd year
students)}\label{literature-review-2nd-year-students}}

The literature review can be on any health economics topic of your
choice, subject to my approval. There is no specific page requirement.
10 double-spaced pages might be a good target, but an
efficiently-written paper could be shorter, while a student wishing to
use the paper as a springboard to a dissertation may choose to write
more. The paper will mostly consist of discussions of prior research,
but should end with discussions of three open questions in the
literature plus a proposed strategy for answering at least one of these
questions. The paper is due by \textbf{Friday, November 13, 2020}.

\hypertarget{draft-paper-3rd-year-students}{%
\subsection{Draft paper (3rd year
students)}\label{draft-paper-3rd-year-students}}

The paper requirement is more extensive for 3rd year students. In this
case, your papers should include not only an extensive literature review
but also a preliminary empirical analysis, including a discussion of the
data you're using, the construction of your sample, your identification
strategy, econometric model and estimator, and some preliminary results.
There is again no specific page requirement, but I expect at least 20
double-spaced pages to appropriately discuss your topic, context, data,
and early results. Your draft paper should end with an outline of
additional analyses you hope to run (i.e., robustness/sensitivity
analyses, a discussion of different mechanisms of interest,
policy-relevant heterogeneities in your estimated effects, etc.). The
paper is due by \textbf{Friday, November 13, 2020}.

\hypertarget{replication}{%
\subsection{Replication}\label{replication}}

You must replicate the empirical analysis of a recent health economics
paper. Your final product should consist of a discussion of the data
sources, steps taken to clean the data and replicate the final analytic
sample, a summary of your model and estimator, and your results. You
must turn in your assignment on \emph{GitHub}, including your code files
(Stata, R, Python, SAS, etc.), all figures/tables, and of course the
final paper and presentation.

Please organize your folders in a useful way. The way I organize things
(though certainly not the only way to do it) is to keep a folder for
each new project and named accordingly. I typically have the following
subfolders:

\begin{itemize}
\item
  Data: This is where I keep the raw data files and any additional data
  files I create as part of my analysis. I also keep a ``Research Data''
  folder on my computer that has raw data that I access regularly, in
  which case the ``Data'' folder in any given project is just a cleaned
  version of the larger files.
\item
  Analysis: This is where I keep my code files, log files, and results.
  I typically have a subfolder for the log files since I create one
  every time I run a code file as well as a subfolder for results.
\item
  Manuscript: This is where I keep all of my files for the actual paper.
  I usually have a different subfolder for every month or two (whenever
  I create a sufficiently different version of the paper).
\item
  Presentation: This is where I keep my slides and underlying code
  files.
\end{itemize}

It's good to start developing some organization practices that work best
for you. It's extremely easy to forget what you were doing on a project
once you have several things going at once, especially when you wait for
6-8 months after submitting a paper for publication. The last thing you
want is to not be able to replicate your own work!

I will provide a set of papers from which you can choose to replicate,
but you are also free to select your own. You need to get approval from
me on a paper to replicate by \textbf{Friday, September 4}. Final
replications are due no later than \textbf{Friday, December 11, 2020}.

\hypertarget{due-dates}{%
\subsection{Due dates}\label{due-dates}}

This section is just to highlight the most important dates on which an
assignment is due. These dates are repeated within the description of
each assignment above. Note that you can always turn assignments in
prior to the official due date. Late assignments will receive an
automatic 10\% reduction in the grade for each day the assignment is
turned in after the due date.

\begin{itemize}
\tightlist
\item
  \textbf{August 21}: Send me the references for the 2 papers that you
  would like to present throughout the semester. Only one student can
  present each paper, so this is first-come first-served.
\item
  \textbf{September 4}: Send me the name of the paper you'd like to
  replicate for the main class project.
\item
  \textbf{Friday, November 13}: Due date for literature review or draft
  paper.
\item
  \textbf{Friday, December 11}: Due date for the replication.
\item
  Empirical exercises due one week after the end of each module
\end{itemize}

\hypertarget{final-grades}{%
\subsection{Final grades}\label{final-grades}}

\begin{itemize}
\item
  \textbf{5\%} for participation in online discussions
\item
  \textbf{20\%} for empirical exercises (5\% each)
\item
  \textbf{20\%} for presentations of selected papers (10\% each)
\item
  \textbf{25\%} for literature review (2nd year students) or draft paper
  (3rd year students)
\item
  \textbf{30\%} for a replication of a selected paper
\end{itemize}

Letter grades will be assigned at the end of the course based on total
score achieved: (A = 100-93\%, A- = 92.99-90\%, B+ = 89.99-87\%, B =
86.99-83\%, B- = 82.99-80\%, C+ = 79.99-77\%, C = 76.99-73\%, C- =
72.99-70\%, D+ = 69.99-67\%, D = 66.99-60\%, F = 60\% or less)

\newpage

\hypertarget{class-schedule}{%
\section{Class Schedule}\label{class-schedule}}

Below is a preliminary outline of specific topics and readings
throughout the semester. Based on our collective interests, discussions,
and timing, individual papers and order may change somewhat, but any
such changes will be announced in advance. For each class, one of us
will present the ``primary reading,'' followed by a general discussion
of the papers and topic among the class. Where relevant, we will work
through some simulation exercises or derive some theoretical results.
Papers listed as ``supplemental'' represent other relevant papers in
that area that we won't have time to cover directly in class. Days with
** indicate that no one will present a paper on that day. Those days are
reserved for some basic theoretical derivations or presentation of new
empirical methods.

\hypertarget{module-0-kicking-things-off}{%
\subsection{Module 0: Kicking things
off!}\label{module-0-kicking-things-off}}

\hypertarget{class-01-0819-introduction-to-the-economics-of-healthcare}{%
\subsubsection{Class 01, 08/19: Introduction to the economics of
healthcare}\label{class-01-0819-introduction-to-the-economics-of-healthcare}}

\textbf{General references:}

{[}1{]} K. J. Arrow. ``Uncertainty and the Welfare Economics of Medical
Care''. In: \emph{The American Economics Review} 53.5 (1963).

{[}2{]} D. Dranove and M. A. Satterthwaite. ``The industrial
organization of health care markets''. In: \emph{Handbook of health
economics} 1 (2000), pp. 1093-1139.

{[}3{]} W. N. Evans, H. Levy, and K. I. Simon. ``Data watch: Research
data in health economics''. In: \emph{Journal of Economic Perspectives}
14.4 (2000), pp.~203-216.

\hypertarget{module-1-the-hospital-objective-functions-and-financial-incentives}{%
\subsection{Module 1: The hospital objective function(s) and financial
incentives}\label{module-1-the-hospital-objective-functions-and-financial-incentives}}

\hypertarget{class-02-0824-theoretical-models-of-not-for-profits}{%
\subsubsection{Class 02, 08/24: Theoretical models of
not-for-profits**}\label{class-02-0824-theoretical-models-of-not-for-profits}}

\textbf{Primary reading:}

{[}1{]} D. Dranove. ``Pricing by non-profit institutions: the case of
hospital cost-shifting''. In: \emph{Journal of Health Economics} 7.1
(1988), pp.~47-57.

{[}2{]} J. P. Newhouse. ``Toward a theory of nonprofit institutions: An
economic model of a hospital''. In: \emph{American Economic Review}
(1970), pp.~64-74.

{[}3{]} M. V. Pauly. ``Medical staff characteristics and hospital
costs''. In: \emph{Journal of Human Resources} (1978), pp.~77-111.

\textbf{Supplemental:}

{[}1{]} F. A. Sloan. ``Not-for-profit ownership and hospital behavior''.
In: \emph{Handbook of health economics} 1 (2000), pp.~1141-1174.

\hypertarget{class-03-0826-empirical-evidence-on-ownership-type}{%
\subsubsection{Class 03, 08/26: Empirical evidence on ownership
type}\label{class-03-0826-empirical-evidence-on-ownership-type}}

\textbf{Primary reading:}

{[}1{]} E. E. Bayindir. ``Hospital ownership type and treatment
choices''. In: \emph{Journal of health economics} 31.2 (2012),
pp.~359-370.

{[}2{]} M. Duggan. ``Hospital Ownership and Public Medical Spending''.
In: \emph{Quarterly Journal of Economics} 115.4 (2000), pp.~1343-1373.

{[}3{]} J. R. Horwitz and A. Nichols. ``Hospital ownership and medical
services: market mix, spillover effects, and nonprofit objectives''. In:
\emph{Journal of health economics} 28.5 (2009), pp.~924-937.

\textbf{Supplemental:}

{[}1{]} T. Chang and M. Jacobson. \emph{What do Nonprofit Hospitals
Maximize. Evidence from California's Seismic Retrofit Mandate}. Working
Paper. Case Western Reserve University, 2017.

{[}2{]} G. David. ``The convergence between for-profit and nonprofit
hospitals in the United States''. In: \emph{International journal of
health care finance and economics} 9.4 (2009), pp.~403-428.

{[}3{]} D. Deneffe and R. T. Masson. ``What do not-for-profit hospitals
maximize?'' In: \emph{International Journal of Industrial Organization}
20.4 (2002), pp.~461-492.

{[}4{]} M. Duggan. ``Hospital market structure and the behavior of
not-for-profit hospitals''. In: \emph{RAND Journal of Economics} (2002),
pp. 433-446.

{[}5{]} R. G. Frank and D. S. Salkever. ``The supply of charity services
by nonprofit hospitals: motives and market structure''. In: \emph{RAND
journal of economics} (1991), pp.~430-445.

{[}6{]} M. Gaynor and W. B. Vogt. ``Competition among Hospitals''. In:
\emph{RAND Journal of Economics} (2003), pp.~764-785.

{[}7{]} J. Gruber. ``The effect of competitive pressure on charity:
Hospital responses to price shopping in California''. In: \emph{Journal
of Health Economics} 13.2 (1994), pp.~183-211.

{[}8{]} D. P. Kessler, M. B. McClellan, and others. ``The Effects of
Hospital Ownership on Medical Productivity''. In: \emph{RAND Journal of
Economics} 33.3 (2002), pp.~488-506.

{[}9{]} F. A. Sloan, G. A. Picone, D. H. Taylor, et al.~``Hospital
ownership and cost and quality of care: is there a dime's worth of
difference?'' In: \emph{Journal of health economics} 20.1 (2001),
pp.~1-21.

\hypertarget{class-04-0831-response-to-medicare-payment-changes}{%
\subsubsection{Class 04, 08/31: Response to Medicare payment
changes}\label{class-04-0831-response-to-medicare-payment-changes}}

\textbf{Primary reading:}

{[}1{]} L. S. Dafny. ``How Do Hospitals Respond to Price Changes?'' In:
\emph{American Economic Review} 95.5 (2005), pp.~1525-1547.

{[}2{]} D. Dranove and C. Ody. ``Employed for higher pay? How Medicare
facility fees affect hospital employment of physicians''. In:
\emph{American Economics Journal: Economic Policy} 11.4 (2019),
pp.~249-271.

\textbf{Supplemental:}

{[}1{]} A. Cook and S. Averett. ``Do Hospitals Respond to Changing
Incentive Structures? Evidence from Medicare's 2007 DRG Restructuring''.
In: \emph{Journal of Health Economics} (2020), p.~102319.

{[}2{]} S. Desai and J. M. McWilliams. ``Consequences of the 340B drug
pricing program''. In: \emph{New England Journal of Medicine} 378.6
(2018), pp.~539-548.

{[}3{]} E. Silverman and J. Skinner. ``Medicare upcoding and hospital
ownership''. In: \emph{Journal of health economics} 23.2 (2004),
pp.~369-389.

\hypertarget{class-05-0902-hospital-cost-shifting}{%
\subsubsection{Class 05, 09/02: Hospital
cost-shifting}\label{class-05-0902-hospital-cost-shifting}}

\textbf{Primary reading:}

{[}1{]} D. Dranove. ``Pricing by non-profit institutions: the case of
hospital cost-shifting''. In: \emph{Journal of Health Economics} 7.1
(1988), pp.~47-57.

{[}2{]} D. Dranove, C. Garthwaite, and C. Ody. ``How do nonprofits
respond to negative wealth shocks? The impact of the 2008 stock market
collapse on hospitals''. In: \emph{The Rand Journal of Economics} 48.2
(2017), pp. 485-525.

\textbf{Supplemental:}

{[}1{]} M. Darden, I. McCarthy, and E. Barrette. \emph{Who Pays in
Pay-for-Performance? Evidence from Hospital Pricing}. Working Paper
w24304. National Bureau of Economic Research, 2018.

{[}2{]} G. David, R. C. Lindrooth, L. A. Helmchen, et al.~``Do hospitals
cross-subsidize?'' In: \emph{Journal of health economics} 37 (2014), pp.
198-218.

{[}3{]} A. B. Frakt. ``How much do hospitals cost shift? A review of the
evidence''. In: \emph{Milbank Quarterly} 89.1 (2011), pp.~90-130.

{[}4{]} J. W. Hay. ``The impact of public health care financing policies
on private-sector hospital costs''. In: \emph{Journal of Health
Politics, Policy and Law} 7.4 (1983), pp.~945-952.

{[}5{]} V. Y. Wu. ``Hospital cost shifting revisited: new evidence from
the balanced budget act of 1997''. In: \emph{International journal of
health care finance and economics} 10.1 (2010), pp.~61-83.

{[}6{]} J. Zwanziger, G. A. Melnick, A. Bamezai, et al.~``Can cost
shifting continue in a price competitive environment?'' In: \emph{Health
Economics} 9.3 (2000), pp.~211-226.

\hypertarget{class-06-0907-response-to-insurance-policies}{%
\subsubsection{Class 06, 09/07: Response to insurance
policies}\label{class-06-0907-response-to-insurance-policies}}

\textbf{Primary reading:}

{[}1{]} M. Batty and B. Ippolito. ``Financial incentives, hospital care,
and health outcomes: Evidence from fair pricing laws''. In:
\emph{American Economic Journal: Economic Policy} 9.2 (2017), pp.~28-56.

{[}2{]} P. J. Eliason, P. L. Grieco, R. C. McDevitt, et al.~``Strategic
patient discharge: The case of long-term care hospitals''. In:
\emph{American Economic Review} 108.11 (2018), pp.~3232-65.

{[}3{]} M. Geruso and T. Layton. ``Upcoding: Evidence from Medicare on
squishy risk adjustment''. In: \emph{Journal of Political Economy} 128.3
(2020), pp.~984-1026.

\textbf{Supplemental:}

{[}1{]} G. Gowrisankaran, K. A. Joiner, and J. Lin. \emph{How do
Hospitals Respond to Payment Incentives?} Tech. rep. National Bureau of
Economic Research, 2019.

{[}2{]} A. Lee. ``How do hospitals respond to managed care? Evidence
from at-risk newborns''. In: \emph{Journal of Public Economics} 184
(2020), p. 104130.

\hypertarget{class-07-0909-working-with-data}{%
\subsubsection{Class 07, 09/09: Working with
data}\label{class-07-0909-working-with-data}}

\textbf{Causal inference with panel data}

\begin{itemize}
\tightlist
\item
  Difference-in-differences and parallel trends
\item
  Event studies and time-varying treatment effects
\item
  Inference with few treated groups
\end{itemize}

\textbf{Applications and data}

\begin{itemize}
\tightlist
\item
  Provider of Services files and Inpatient Prospective Payment System
  Final Rule files
\item
  Replicate parts of Dafny (2005)
\end{itemize}

\hypertarget{module-2-physician-agency-and-treatment-decisions}{%
\subsection{Module 2: Physician agency and treatment
decisions}\label{module-2-physician-agency-and-treatment-decisions}}

\hypertarget{class-08-0914-the-agency-problem-in-healthcare}{%
\subsubsection{Class 08, 09/14: The agency problem in
healthcare**}\label{class-08-0914-the-agency-problem-in-healthcare}}

\textbf{Primary reading:}

{[}1{]} T. G. McGuire. ``Physician agency''. In: \emph{Handbook of
health economics} 1 (2000), pp.~461-536.

\textbf{Supplemental:}

{[}1{]} A. Chandra, A. Jena, and J. Skinner. ``The Pragmatist's Guide to
Comparative Effectiveness Research''. In: \emph{The Journal of Economic
Perspectives} 25.2 (2011), pp.~27-46.

{[}2{]} S. J. Grossman and O. D. Hart. ``An analysis of the
principal-agent problem''. In: \emph{Econometrica: Journal of the
Econometric Society} (1983), pp.~7-45.

\hypertarget{class-09-0916-supply-side-variation-in-healthcare}{%
\subsubsection{Class 09, 09/16: Supply-side variation in
healthcare}\label{class-09-0916-supply-side-variation-in-healthcare}}

\textbf{Primary reading:}

{[}1{]} A. Finkelstein, M. Gentzkow, and H. Williams. ``Sources of
geographic variation in health care: Evidence from patient migration''.
In: \emph{The quarterly journal of economics} 131.4 (2016),
pp.~1681-1726.

{[}2{]} D. Molitor. ``The evolution of physician practice styles:
evidence from cardiologist migration''. In: \emph{American Economic
Journal: Economic Policy} 10.1 (2018), pp.~326-56.

\hypertarget{class-10-0921-physician-response-to-payment-changes}{%
\subsubsection{Class 10, 09/21: Physician response to payment
changes}\label{class-10-0921-physician-response-to-payment-changes}}

\textbf{Primary reading:}

{[}1{]} J. Clemens and J. D. Gottlieb. ``Do Physicians' Financial
Incentives Affect Medical Treatment and Patient Health?'' In:
\emph{American Economic Review} 104.4 (2014), pp.~1320-1349.

{[}2{]} J. Gruber and M. Owings. ``Physician financial incentives and
cesarean section delivery''. In: \emph{The RAND Journal of Economics}
27.1 (1996), pp.~99-123.

\textbf{Supplemental:}

{[}1{]} L. C. Baker, M. K. Bundorf, and D. P. Kessler. ``The Effect of
Hospital/Physician Integration on Hospital Choice''. In: \emph{Journal
of Health Economics} 50 (2016), pp.~1-8.

{[}2{]} W. Beckert. ``Choice in the presence of experts: The role of
general practitioners in patients' hospital choice''. In: \emph{Journal
of Health Economics} 60 (2018), pp.~98 - 117. ISSN: 0167-6296. DOI:
\url{https://doi.org/10.1016/j.jhealeco.2018.06.003}. \textless URL:
\url{http://www.sciencedirect.com/science/article/pii/S0167629617302722}\textgreater.

{[}3{]} M. Gaynor, C. Propper, and S. Seiler. ``Free to Choose? Reform,
Choice, and Consideration Sets in the English National Health Service''.
In: \emph{American Economic Review} 106.11 (Nov.~2016), pp.~3521-57.
DOI: 10.1257/aer.20121532. \textless URL:
\url{http://www.aeaweb.org/articles?id=10.1257/aer.20121532}\textgreater.

\hypertarget{class-11-0923-physician-spillovers}{%
\subsubsection{Class 11, 09/23: Physician
spillovers}\label{class-11-0923-physician-spillovers}}

\textbf{Primary reading:}

{[}1{]} K. Baicker, M. E. Chernew, and J. A. Robbins. ``The spillover
effects of Medicare managed care: Medicare Advantage and hospital
utilization''. In: \emph{Journal of health economics} 32.6 (2013), pp.
1289-1300.

{[}2{]} J. Clemens and J. D. Gottlieb. ``In the shadow of a giant:
Medicare's influence on private physician payments''. In: \emph{Journal
of Political Economy} 125.1 (2017), pp.~1-39.

\textbf{Supplemental:}

{[}1{]} L. C. Baker. ``Managed care spillover effects''. In:
\emph{Annual review of public health} 24.1 (2003), pp.~435-456.

{[}2{]} A. Chandra and D. O. Staiger. ``Productivity spillovers in
health care: evidence from the treatment of heart attacks''. In:
\emph{Journal of Political Economy} 115.1 (2007), pp.~103-140.

{[}3{]} M. Chernew, K. Baicker, and C. Martin. \emph{Spillovers in
Health Care Markets: Implications for Current Law Projections.} Report.
Centers for Medicare and Medicaid Services, 2010.

{[}4{]} R. G. Frank and R. J. Zeckhauser. ``Custom-made versus
ready-to-wear treatments: Behavioral propensities in physicians'
choices''. In: \emph{Journal of health economics} 26.6 (2007),
pp.~1101-1127.

{[}5{]} S. Glied and J. G. Zivin. ``How do doctors behave when some (but
not all) of their patients are in managed care?'' In: \emph{Journal of
Health Economics} 21.2 (2002), pp.~337-353.

{[}6{]} B. E. Landon. ``Tipping the scale: The norms hypothesis and
primary care physician behavior''. In: \emph{New England Journal of
Medicine} 376.9 (2017), pp.~810-811.

{[}7{]} T. G. McGuire and M. V. Pauly. ``Physician response to fee
changes with multiple payers''. In: \emph{Journal of health economics}
10.4 (1991), pp.~385-410.

{[}8{]} J. P. Newhouse and M. S. Marquis. ``The norms hypothesis and the
demand for medical care''. In: \emph{Journal of Human Resources} (1978),
pp. 159-182.

\hypertarget{class-12-0928-working-with-data}{%
\subsubsection{Class 12, 09/28: Working with
data}\label{class-12-0928-working-with-data}}

\textbf{Instrumental variables}

\begin{itemize}
\tightlist
\item
  What is a good instrument?
\item
  What are we estimating with IV?
\item
  Specification tests and current robustness checks
\end{itemize}

\textbf{Applications and data}

\begin{itemize}
\tightlist
\item
  Medicare Advantage data and Area Health Resources File
\item
  Replicate elements of Baicker, Chernew, and Robbins (2013)
\end{itemize}

\hypertarget{module-3-information-disclosure-and-patient-decision-making}{%
\subsection{Module 3: Information disclosure and patient decision
making}\label{module-3-information-disclosure-and-patient-decision-making}}

\hypertarget{class-13-0930-poor-decisions-in-healthcare-and-health-insurance}{%
\subsubsection{Class 13, 09/30: Poor decisions in healthcare and health
insurance}\label{class-13-0930-poor-decisions-in-healthcare-and-health-insurance}}

\textbf{Primary reading:}

{[}1{]} M. Chernew, Z. Cooper, E. Larsen-Hallock, et al.~\emph{Are
Health Care Services Shoppable? Evidence from the Consumption of
Lower-Limb MRI Scans}. Working Paper 24869. National Bureau of Economic
Research, Jul. 2018. \textless URL:
\url{http://www.nber.org/papers/w24869}\textgreater.

{[}2{]} J. D. Ketcham, C. Lucarelli, E. J. Miravete, et al.~``Sinking,
swimming, or learning to swim in Medicare Part D''. In: \emph{American
Economic Review} 102.6 (2012), pp.~2639-2673.

\textbf{Supplemental:}

{[}1{]} J. Abaluck and J. Gruber. ``Heterogeneity in choice
inconsistencies among the elderly: evidence from prescription drug plan
choice''. In: \emph{American Economic Review} 101.3 (2011), pp.~377-381.

{[}2{]} J. Farrell and P. Klemperer. ``Coordination and lock-in:
Competition with switching costs and network effects''. In:
\emph{Handbook of industrial organization} 3 (2007), pp.~1967-2072.

\hypertarget{class-14-1005-inertia-in-health-insurance-choice}{%
\subsubsection{Class 14, 10/05: Inertia in health insurance
choice}\label{class-14-1005-inertia-in-health-insurance-choice}}

\textbf{Primary reading:}

{[}1{]} K. M. Ericson. ``Consumer inertia and firm pricing in the
Medicare Part D prescription drug insurance exchange''. In:
\emph{American Economic Journal: Economic Policy} 6.1 (2014), pp.~38-64.

{[}2{]} B. R. Handel. ``Adverse selection and inertia in health
insurance markets: When nudging hurts''. In: \emph{American Economic
Review} 103.7 (2013), pp.~2643-2682.

\hypertarget{class-15-1007-effects-of-quality-disclosure}{%
\subsubsection{Class 15, 10/07: Effects of quality
disclosure}\label{class-15-1007-effects-of-quality-disclosure}}

\textbf{Primary reading:}

{[}1{]} L. Dafny and D. Dranove. ``Do report cards tell consumers
anything they don't already know? The case of Medicare HMOs''. In:
\emph{The Rand journal of economics} 39.3 (2008), pp.~790-821.

{[}2{]} D. Dranove, D. Kessler, M. McClellan, et al.~``Is More
Information Better? The Effects of Report Cards on Health Care
Providers''. In: \emph{The Journal of Political Economy} 111.3 (2003),
pp.~555-588.

\textbf{Supplemental:}

{[}1{]} M. Chernew, D. Scanlon, and others. ``Health plan report cards
and insurance choice.'' In: \emph{Inquiry: a journal of medical care
organization, provision and financing} 35.1 (1998), p.~9.

{[}2{]} M. Darden and I. McCarthy. ``The Star Treatment: Estimating the
Impact of Star Ratings on Medicare Advantage Enrollments''. In:
\emph{Journal of Human Resources} 50.4 (2015), pp.~980-1008.

{[}3{]} D. Dranove and G. Z. Jin. ``Quality Disclosure and
Certification: Theory and Practice''. In: \emph{Journal of Economic
Literature} 48.4 (2010), pp.~935-963.

{[}4{]} G. Jin and A. Sorensen. ``Information and consumer choice: the
value of publicized health plan ratings''. In: \emph{Journal of Health
Economics} 25.2 (2006), pp.~248-275.

{[}5{]} I. McCarthy and M. Darden. ``Supply-side Responses to Public
Quality Ratings: Evidence from Medicare Advantage''. In: \emph{American
Journal of Health Economics} 3.2 (2017), pp.~140-164.

{[}6{]} R. O. Reid, P. Deb, B. L. Howell, et al.~``Association Between
Medicare Advantage Plan Star Ratings and EnrollmentStar Ratings for
Medicare Advantage Plan''. In: \emph{JAMA} 309.3 (2013), pp.~267-274.

{[}7{]} D. Scanlon, M. Chernew, C. McLaughlin, et al.~``The impact of
health plan report cards on managed care enrollment''. In: \emph{Journal
of Health Economics} 21.1 (2002), pp.~19-41.

{[}8{]} G. Wedig and M. Tai-Seale. ``The effect of report cards on
consumer choice in the health insurance market''. In: \emph{Journal of
Health Economics} 21.6 (2002), pp.~1031-1048.

\hypertarget{class-16-1012-effects-of-price-disclosure}{%
\subsubsection{Class 16, 10/12: Effects of price
disclosure}\label{class-16-1012-effects-of-price-disclosure}}

\textbf{Primary reading:}

{[}1{]} M. Grennan and A. Swanson. ``Transparency and negotiated prices:
The value of information in hospital-supplier bargaining''. In:
\emph{Journal of Political Economy} 128.4 (2020), pp.~1234-1268.

\hypertarget{class-17-1014-working-with-data}{%
\subsubsection{Class 17, 10/14: Working with
data}\label{class-17-1014-working-with-data}}

\textbf{Regression discontinuity}

\begin{itemize}
\tightlist
\item
  Basics of strict and fuzzy RD
\item
  Specification tests
\item
  Bin and bandwidth selection
\end{itemize}

\textbf{Applications and data}

\begin{itemize}
\tightlist
\item
  Medicare Part D Data
\item
  Replicate Ericson (2014)
\end{itemize}

\hypertarget{module-4-healthcare-competition-and-industrial-organization}{%
\subsection{Module 4: Healthcare competition and industrial
organization}\label{module-4-healthcare-competition-and-industrial-organization}}

\hypertarget{class-18-1019-competition-in-price-and-quality}{%
\subsubsection{Class 18, 10/19: Competition in price and
quality**}\label{class-18-1019-competition-in-price-and-quality}}

\textbf{Primary reading:}

{[}1{]} D. Dranove and M. A. Satterthwaite. ``Monopolistic competition
when price and quality are imperfectly observable''. In: \emph{The RAND
Journal of Economics} (1992), pp.~518-534.

{[}2{]} Y. Fan. ``Ownership consolidation and product characteristics: A
study of the US daily newspaper market''. In: \emph{American Economic
Review} 103.5 (2013), pp.~1598-1628.

{[}3{]} M. Lewis and K. Pflum. ``Competition and Quality Choice in
Hospital Markets''. In: \emph{Working Paper} (2016).

\textbf{Supplemental:}

{[}1{]} J. D. Dana Jr and Y. Fong. ``Product quality, reputation, and
market structure''. In: \emph{International Economic Review} 52.4
(2011), pp. 1059-1076.

{[}2{]} J. Hausman, G. Leonard, and J. D. Zona. ``Competitive analysis
with differenciated products''. In: \emph{Annales d'Economie et de
Statistique} (1994), pp.~159-180.

{[}3{]} M. Mussa and S. Rosen. ``Monopoly and product quality''. In:
\emph{Journal of Economic theory} 18.2 (1978), pp.~301-317.

{[}4{]} A. Spence. ``Monopoly, Quality, and Regulation''. In: \emph{Bell
Journal of Economics} 6.2 (1975), pp.~417-429.

\hypertarget{class-19-1021-measuring-hospital-market-power}{%
\subsubsection{Class 19, 10/21: Measuring hospital market
power}\label{class-19-1021-measuring-hospital-market-power}}

\textbf{Primary reading:}

{[}1{]} D. Dranove and C. Ody. ``Evolving Measures of Provider Market
Power''. In: \emph{American Journal of Health Economics} 2.2 (2016), pp.
145-160.

\textbf{Supplemental:}

{[}1{]} C. Capps, D. Dranove, and M. Satterthwaite. ``Competition and
market power in option demand markets''. In: \emph{RAND Journal of
Economics} (2003), pp.~737-763.

{[}2{]} D. Dranove and M. Shanley. ``A note on the relational aspects of
hospital market definitions''. In: \emph{Journal of health economics}
8.4 (1990), pp.~473-478.

\hypertarget{class-20-1026-evidence-from-structure-conduct-performance-studies}{%
\subsubsection{Class 20, 10/26: Evidence from
``structure-conduct-performance''
studies}\label{class-20-1026-evidence-from-structure-conduct-performance-studies}}

\textbf{Primary reading:}

{[}1{]} Z. Cooper, S. V. Craig, M. Gaynor, et al.~\emph{The price ain't
right? Hospital prices and health spending on the privately insured}.
Working Paper. National Bureau of Economic Research, 2017.

{[}2{]} D. Kessler and M. McClellan. ``Is hospital competition socially
wasteful?'' In: \emph{Quarterly Journal of Economics} 2.115 (2000), pp.
577-615.

\textbf{Supplemental:}

{[}1{]} D. Dranove, M. Shanley, and C. Simon. ``Is hospital competition
wasteful?'' In: \emph{The RAND journal of economics} (1992),
pp.~247-262.

{[}2{]} E. B. Keeler, G. Melnick, and J. Zwanziger. ``The changing
effects of competition on non-profit and for-profit hospital pricing
behavior''. In: \emph{Journal of health economics} 18.1 (1999),
pp.~69-86.

{[}3{]} W. J. Lynk. ``Nonprofit hospital mergers and the exercise of
market power''. In: \emph{The Journal of Law and Economics} 38.2 (1995),
pp. 437-461.

\hypertarget{class-21-1028-competition-in-a-bargaining-model}{%
\subsubsection{Class 21, 10/28: Competition in a bargaining
model}\label{class-21-1028-competition-in-a-bargaining-model}}

\textbf{Primary reading:}

{[}1{]} K. Ho and R. S. Lee. ``Insurer competition in health care
markets''. In: \emph{Econometrica} 85.2 (2017), pp.~379-417.

\textbf{Supplemental:}

{[}1{]} J. Abraham, M. Gaynor, and W. B. Vogt. ``Entry and Competition
in Local Hospital Markets''. In: \emph{The Journal of Industrial
Economics} 55.2 (2007), pp.~265-288.

{[}2{]} T. F. Bresnahan and P. C. Reiss. ``Entry and competition in
concentrated markets''. In: \emph{Journal of Political Economy} (1991),
pp. 977-1009.

{[}3{]} G. Gowrisankaran, A. Nevo, and R. Town. ``Mergers When Prices
Are Negotiated: Evidence from the Hospital Industry''. In:
\emph{American Economic Review} 105.1 (2015), pp.~172-203.

{[}4{]} M. Lewis and K. Pflum. ``Diagnosing Hospital System Bargaining
Power in Managed Care Networks''. In: \emph{American Economic Journal:
Economic Policy} 7.1 (2015), pp.~243-274.

{[}5{]} P. Reiss, F. Wolak, and E. Learner. ``Handbook of
econometrics''. In: \emph{Structural econometric modeling: Rationales
and examples from industrial organization} 6A (2007).

\hypertarget{class-22-1102-evidence-from-mergers-and-closures}{%
\subsubsection{Class 22, 11/02: Evidence from mergers and
closures}\label{class-22-1102-evidence-from-mergers-and-closures}}

\textbf{Primary reading:}

{[}1{]} L. Dafny. ``Estimation and Identification of Merger Effects: An
Application to Hospital Mergers''. In: \emph{Journal of Law and
Economics} 52.3 (2009), pp.~523-550.

\textbf{Supplemental:}

{[}1{]} M. Gaynor and R. Town. \emph{The impact of hospital
consolidation - Update}. Policy Brief. Robert Wood Johnson Foundation,
2012.

{[}2{]} M. Gaynor and W. B. Vogt. ``Competition among Hospitals''. In:
\emph{RAND Journal of Economics} (2003), pp.~764-785.

{[}3{]} R. C. Lindrooth, A. T. L. Sasso, and G. J. Bazzoli. ``The effect
of urban hospital closure on markets''. In: \emph{Journal of Health
Economics} 22.5 (2003), pp.~691-712.

{[}4{]} M. G. Vita and S. Sacher. ``The competitive effects of
not-for-profit hospital mergers: a case study''. In: \emph{The Journal
of Industrial Economics} 49.1 (2001), pp.~63-84.

\hypertarget{class-23-1104-evidence-from-out-of-market-mergers}{%
\subsubsection{Class 23, 11/04: Evidence from out-of-market
mergers}\label{class-23-1104-evidence-from-out-of-market-mergers}}

\textbf{Primary reading:}

{[}1{]} L. Dafny, K. Ho, and R. S. Lee. ``The price effects of
cross-market mergers: theory and evidence from the hospital industry''.
In: \emph{The RAND Journal of Economics} 50.2 (2019), pp.~286-325.

{[}2{]} M. Schmitt. ``Multimarket Contact in the Hospital Industry''.
In: \emph{American Economic Journal: Economic Policy} 10.3 (2018),
pp.~361-87.

\hypertarget{class-24-1109-mergers-and-acquisitions-in-dialysis}{%
\subsubsection{Class 24, 11/09: Mergers and acquisitions in
dialysis}\label{class-24-1109-mergers-and-acquisitions-in-dialysis}}

\textbf{Primary reading:}

{[}1{]} P. J. Eliason, B. Heebsh, R. C. McDevitt, et al.~``How
Acquisitions Affect Firm Behavior and Performance: Evidence from the
Dialysis Industry''. In: \emph{The Quarterly Journal of Economics} 135.1
(2020), pp. 221-267.

\hypertarget{class-25-1111-vertical-integration-between-hospitals-and-physicians}{%
\subsubsection{Class 25, 11/11: Vertical integration between hospitals
and
physicians}\label{class-25-1111-vertical-integration-between-hospitals-and-physicians}}

\textbf{Primary reading:}

{[}1{]} T. G. Koch, B. W. Wendling, and N. E. Wilson. ``How vertical
integration affects the quantity and cost of care for Medicare
beneficiaries''. In: \emph{Journal of Health Economics} 52 (2017),
pp.~19-32.

{[}2{]} H. Lin, I. McCarthy, and M. Richards. \emph{How does
hospital-physician integration affect hospital prices and quality of
care?} Working Paper. Indiana University, Kelley School of Business,
2020.

{[}3{]} F. Luco and G. Marshall. ``The Competitive Impact of Vertical
Integration by Multiproduct Firms''. In: \emph{American Economic Review}
(2020).

\textbf{Supplemental:}

{[}1{]} L. C. Baker, M. K. Bundorf, and D. P. Kessler. ``The Effect of
Hospital/Physician Integration on Hospital Choice''. In: \emph{Journal
of Health Economics} 50 (2016), pp.~1-8.

{[}2{]} L. R. Burns and M. V. Pauly. ``Integrated delivery networks: a
detour on the road to integrated health care?'' In: \emph{Health
affairs} 21.4 (2002), pp.~128-143.

{[}3{]} C. Capps, D. Dranove, and C. Ody. ``The effect of hospital
acquisitions of physician practices on prices and spending''. In:
\emph{Journal of health economics} 59 (2018), pp.~139-152.

{[}4{]} F. Ciliberto and D. Dranove. ``The effect of physician-hospital
affiliations on hospital prices in California''. In: \emph{Journal of
Health Economics} 25.1 (2006), pp.~29-38.

{[}5{]} A. E. Cuellar and P. J. Gertler. ``Strategic integration of
hospitals and physicians''. In: \emph{Journal of Health Economics} 25.1
(2006), pp.~1-28.

{[}6{]} R. T. Konetzka, E. A. Stuart, and R. M. Werner. ``The effect of
integration of hospitals and post-acute care providers on Medicare
payment and patient outcomes''. In: \emph{Journal of health economics}
(2018).

{[}7{]} H. T. Neprash, M. E. Chernew, A. L. Hicks, et al.~``Association
of financial integration between physicians and hospitals with
commercial health care prices''. In: \emph{JAMA internal medicine}
175.12 (2015), pp. 1932-1939.

\hypertarget{class-26-1116-working-with-data}{%
\subsubsection{Class 26, 11/16: Working with
data}\label{class-26-1116-working-with-data}}

\hypertarget{module-5-health-insurance-markets-likely-wont-get-to-this}{%
\subsection{Module 5: Health insurance markets (likely won't get to
this)}\label{module-5-health-insurance-markets-likely-wont-get-to-this}}

\hypertarget{insurance-markets-and-competitiveness}{%
\subsubsection{Insurance markets and
competitiveness}\label{insurance-markets-and-competitiveness}}

\textbf{Primary reading:}

{[}1{]} M. K. Bundorf, J. Levin, and N. Mahoney. ``Pricing and welfare
in health plan choice''. In: \emph{American Economic Review} 102.7
(2012), pp. 3214-3248.

{[}2{]} L. Dafny. ``Are Health Insurance Markets Competitive?'' In:
\emph{American Economic Review} 100.4 (2010), pp.~1399-1431.

{[}3{]} L. Dafny, M. Duggan, and S. Ramanarayanan. ``Paying a Premium on
Your Premium? Consolidation in the US Health Insurance Industry''. In:
\emph{American Economic Review} 102.2 (2012), pp.~1161-1185.

\textbf{Supplemental:}

{[}1{]} L. Einav, A. Finkelstein, and M. R. Cullen. ``Estimating Welfare
in Insurance Markets Using Variation in Prices''. In: \emph{Quarterly
Journal of Economics} 125.3 (2010), pp.~877-921.

{[}2{]} L. Einav, A. Finkelstein, and J. Levin. ``Beyond Testing:
Empirical Models of Insurance Markets''. In: \emph{Annu. Rev.~Econ.} 2.1
(2010), pp. 311-336.

{[}3{]} A. Starc. ``Insurer pricing and consumer welfare: Evidence from
medigap''. In: \emph{The RAND Journal of Economics} 45.1 (2014),
pp.~198-220.

{[}4{]} R. Town and S. Liu. ``The welfare impact of Medicare HMOs''. In:
\emph{RAND Journal of Economics} (2003), pp.~719-736.

\hypertarget{adverse-selection}{%
\subsubsection{Adverse selection}\label{adverse-selection}}

\textbf{Primary reading:}

{[}1{]} D. M. Cutler and S. J. Reber. ``Paying for Health Insurance: The
Trade-Off between Competition and Adverse Selection''. In:
\emph{Quarterly Journal of Economics} 113.2 (1998), pp.~433-466.

{[}2{]} F. Decarolis and A. Guglielmo. ``Insurers' response to selection
risk: Evidence from Medicare enrollment reforms''. In: \emph{Journal of
health economics} 56 (2017), pp.~383-396.

\textbf{Supplemental:}

{[}1{]} G. Ackerloff. ``The market for lemons: Quality uncertainty and
the market mechanism''. In: \emph{Quarterly Journal of Economics} 84.3
(1970), pp.~488-500.

{[}2{]} L. Einav and A. Finkelstein. ``Selection in Insurance Markets:
Theory and Empirics in Pictures''. In: \emph{The Journal of Economic
Perspectives} 25.1 (2011), pp.~115-138.

{[}3{]} R. G. Frank, J. Glazer, and T. G. McGuire. ``Measuring adverse
selection in managed health care''. In: \emph{Journal of Health
Economics} 19.6 (2000), pp.~829-854.

{[}4{]} M. Rothschild and J. Stiglitz. ``Equilibrium in competitive
insurance markets: An essay on the economics of imperfect information''.
In: \emph{Quarterly Journal of Economics} (1976), pp.~629-649.

\hypertarget{moral-hazard}{%
\subsubsection{Moral hazard}\label{moral-hazard}}

\textbf{Primary reading:}

{[}1{]} L. Einav, A. Finkelstein, S. P. Ryan, et al.~``Selection on
moral hazard in health insurance''. In: \emph{American Economic Review}
103.1 (2013), pp.~178-219.

\textbf{Supplemental:}

{[}1{]} L. Einav and A. Finkelstein. \emph{Moral Hazard in Health
Insurance: What We Know and How We Know It}. Tech. rep. 24055. National
Bureau of Economic Research, 2017.

{[}2{]} A. Finkelstein. \emph{Moral hazard in health insurance}.
Columbia University Press, 2014.

{[}3{]} W. G. Manning and M. S. Marquis. ``Health insurance: the
tradeoff between risk pooling and moral hazard''. In: \emph{Journal of
health economics} 15.5 (1996), pp.~609-639.

\hypertarget{managed-competition}{%
\subsubsection{Managed competition}\label{managed-competition}}

\textbf{Primary reading:}

{[}1{]} V. Curto, L. Einav, J. Levin, et al.~\emph{Can Health Insurance
Competition Work? Evidence from Medicare Advantage}. Working Paper.
Department of Economics, Stanford University, 2015.

{[}2{]} L. Einav and J. Levin. ``Managed competition in health
insurance''. In: \emph{Journal of the European Economic Association}
13.6 (2015), pp. 998-1021.

\textbf{Supplemental:}

{[}1{]} M. Cabral, M. Geruso, and N. Mahoney. \emph{Does privatized
health insurance benefit patients or producers? Evidence from Medicare
Advantage}. Tech. rep. 20470. National Bureau of Economic Research,
2014.

{[}2{]} M. Duggan, A. Starc, and B. Vabson. ``Who benefits when the
government pays more? Pass-through in the Medicare Advantage program''.
In: \emph{Journal of Public Economics} 141 (2016), pp.~50-67.

{[}3{]} D. Pelech. ``Paying more for less? Insurer competition and
health plan generosity in the Medicare Advantage program''. In:
\emph{Journal of Health Economics} 61 (2018), pp.~77-92.

{[}4{]} Z. Song, M. B. Landrum, and M. E. Chernew. ``Competitive bidding
in Medicare Advantage: Effect of benchmark changes on plan bids''. In:
\emph{Journal of health economics} 32.6 (2013), pp.~1301-1312.

{[}5{]} K. Stockley, T. McGuire, C. Afendulis, et al.~\emph{Premium
Transparency in the Medicare Advantage Market: Implications for
Premiums, Benefits, and Efficiency}. Working Paper. National Bureau of
Economic Research, 2014.




\end{document}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother